\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-05-14 doc/sw-system.tex
Addressed Anu's final queries, closes #210. Papers now look consistent, closed #207

We describe \SW, a novel gravitational lens discovery service that yields
samples of high purity and completeness through crowd-sourced visual inspection.
Carefully produced colour composite images are displayed to volunteers via a
web-based classification interface, which records their estimates of the positions of
candidate lensed features. Simulated lenses, and expert-classified images which
lack lenses, are inserted into the image stream at random intervals; this
training set is used to give the volunteers instantaneous feedback on their performance, as
well as to calibrate a model of the system that provides
dynamical updates to the probability that a classified image contains a lens. Low probability systems are
retired from the site periodically, concentrating the sample towards a set of lens
candidates. Having divided 160 square degrees of Canada-France-Hawaii Telescope
Legacy Survey (\cfhtls) imaging into some 430,000 overlapping 82 by 82 arcsecond
tiles and displaying them on the site, we were joined by around 37,000
volunteers who contributed 11~million image classifications over the course of 8
months. This \StageOne search reduced the sample to 3368 images containing candidates; these were then
refined in \StageTwo to yield a sample that we expect to be over 90\% complete and 30\% pure, based on our analysis of the volunteers performance on training images (refer for \Sref{sec:design:stages} for more information on the two stage strategy) .
We comment on the scalability of the \SW system to the wide field survey era,
based on our finding that searches of $10^4$ images can be performed by a
crowd of $10^4$ volunteers in 6 days.
\end{abstract}

@@

How can we scale these lens searches up to imaging surveys that cover a hundred
times the sky area, such as the almost-all sky surveys planned with the Large
Synoptic Survey Telescope (LSST) and
Euclid?  There are two approaches to detecting lenses in imaging surveys. The
first one is robotic: automated analysis of the object catalogs and the survey
images. The candidate samples produced by these methods have, to date, not been
of high purity  \citep[see
e.g.][]{MarshallEtal2009,MoreEtal2012,GavazziEtal2014}, with visual inspection
by teams of humans still required to narrow down the robotically-generated
samples. In this approach, the image data may or may not be explicitly modelled
by the robots as if it contained a gravitational lens, but the visual inspection
can be thought of as a ``mental modeling'' step.  An inspector who classifies an
object as a lens candidate does so because the features in the image that they
see can be explained by a model, contained in their brain, of what gravitational
lenses do.  The second approach simply cuts out the robot middleman:
\citet{MoustakasEtal2007,FaureEtal2008,Jackson2008} and \citet{PawaseEtal2014}
all performed successful visual searches for lenses in HST imaging.

@@

We refer to this alternative calculation as ``offline'' analysis, because it
does not need to be carried out one classification at a time (and hence can be
done at any time after the data is collected). Note that in this offline
inference, the effect will be that of applying the time-averaged confusion
matrices to each classification,  rather than a set that evolves as the agents
(and in the real world, the volunteers) learn. This will mean  that the early
classifications will effectively not be downweighted as a result of the agent's
ignorance. On the other hand, this ignorance could also have provided some
conservatism, reducing the noise due to early classifications if they are
unreliable. Whether or not an offline analysis out-performs an online one is a
matter for experiment.

@@

How fast does the \SW crowd classify subjects? Each software agent records the timestamp
of each classification its volunteer makes; by measuring the time lags between
successive classifications, we can make estimates of the crowd's classification
and contribution speed. We plot the former quantity for both \StageOne and
\StageTwo in \Fref{fig:results:speed}, normalizing the speed and time axes to
their respective totals. The fractional classification rates
in each stage of the survey fall off in approximately the same way, despite the
factor of nearly 20 difference in crowd size. \StageOne (consisting of
$\sim430,000$ subjects) was completed in around 5000 hours (by some 37,000
participants), with classification rates in the first few days averaging at
$\sim10^4$ per hour. This was stimulated by various forms of advertising (press
releases, and emails to registered Zooniverse users). The asymptotic
classification rate was  around ten times lower, at $\sim10^3$ per hour.
Interestingly, the average skill was found to be approximately constant over the
lifetime of \StageOne, leading to a contribution rate that tracks closely the
classification rate. (A 10\% increase in average skill was seen over the first
40 days, and a decrease of the same amount in the last 40 days -- not enough to
cause a significant difference between classification and contribution rate
behaviour, but perhaps reflecting a learning period at the start and then a
decrease
in participation later on.) \StageTwo (3400 subjects) was completed in around 600
hours by around 2000 participants, with classification rates starting at
$~\sim2000$ per hour and decaying to $\sim100$ per hour.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
\begin{center}
\caption{Total crowd and subject sample properties from the \cfhtls project.}
\label{tab:crowd:contributions}
\begin{tabular}{cccccccc}
  \hline
  \hline {Stage} & Subjects & Contribution                          & Agents & Skill      & Classifications          & Candidates & Information \\
                           & $\Ns$    & $\sum_k^{\Nv} \contribution_k$ (bits) & $\Nv$  & $\sum_k^{\Nv} \skill_k$ (bits) & $\sum_k^{\Nv} \thiseffort$ & $\Ncands$  & $\sum_j^{\Ns}\sum_k^{\Nv} \information_{j,k}$ (bits) \\
            \hline
                      1    & $427064$ & $1292016.3$ & $36982$ & $1471.9$ & $10802125$ & $3368$ & $91122.6$ \\
                      2    & $3679$   &   $21895.8$ &  $1964$ &  $102.4$ &   $224745$ &   $89$ &  $1640.4$ \\
  \hline \hline
\end{tabular}
\medskip\\
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $91-93\%$ and $<1\%$
(Stages I and II). This FPR translates to an approximate purity of 30\%. We find a
higher purity at \StageTwo (perhaps as a result of the least visible candidates already
being discarded): here, 100\% purity was reached at just under $90\%$
completeness. Because \SW is a supervised learning system, these numbers should
be taken to be upper limits to what we should expect for real lenses (which have
not been selected for a high visibility training set).

>>>>

We describe \SW, a novel gravitational lens discovery service that yields
samples of high purity and completeness through crowd-sourced visual inspection.
Carefully produced colour composite images are displayed to volunteers via a
web-based classification interface, which records their estimates of the positions of
candidate lensed features. Simulated lenses, and expert-classified images which
lack lenses, are inserted into the image stream at random intervals; this
training set is used to give the volunteers instantaneous feedback on their performance, as
well as to calibrate a model of the system that provides
dynamical updates to the probability that a classified image contains a lens. Low probability systems are
retired from the site periodically, concentrating the sample towards a set of lens
candidates. Having divided 160 square degrees of Canada-France-Hawaii Telescope
Legacy Survey (\cfhtls) imaging into some 430,000 overlapping 82 by 82 arcsecond
tiles and displaying them on the site, we were joined by around 37,000
volunteers who contributed 11~million image classifications over the course of 8
months. This \StageOne search reduced the sample to 3381 images containing candidates; these were then
refined in \StageTwo to yield a sample that we expect to be over 90\% complete and 30\% pure, based on our analysis of the volunteers performance on training images (refer for \Sref{sec:design:stages} for more information on the two stage strategy) .
We comment on the scalability of the \SW system to the wide field survey era,
based on our finding that searches of $10^4$ images can be performed by a
crowd of $10^4$ volunteers in 6 days.
\end{abstract}

@@

How can we scale these lens searches up to imaging surveys that cover a hundred
times the sky area, such as the almost-all sky surveys planned with the Large
Synoptic Survey Telescope (LSST) and
Euclid?  There are two approaches to detecting lenses in imaging surveys. The
first one is robotic: automated analysis of the object catalogs and the survey
images. The candidate samples produced by these methods have, to date, only
reached purities of 1-10\%, with visual inspection
by teams of humans still required to reduce the robotically-generated
samples by factors of 10-100 \citep[see
e.g.][]{MarshallEtal2009,MoreEtal2012,GavazziEtal2014}.
In this approach, the image data may or may not be explicitly modelled
by the robots as if it contained a gravitational lens, but the visual inspection
can be thought of as a ``mental modeling'' step.  An inspector who classifies an
object as a lens candidate does so because the features in the image that they
see can be explained by a model, contained in their brain, of what gravitational
lenses do.  The second approach simply cuts out the robot middleman:
\citet{MoustakasEtal2007,FaureEtal2008,Jackson2008} and \citet{PawaseEtal2014}
all performed successful visual searches for lenses in HST imaging.

@@

We refer to this alternative calculation as ``offline'' analysis, because it
does not need to be carried out one classification at a time (and hence can be
done at any time after the data is collected). Note that in this offline
inference, the effect will be that of applying the time-averaged confusion
matrices to each classification,  rather than a set that evolves as the agents
(and in the real world, the volunteers) learn. This will mean  that the early
classifications will effectively not be downweighted as a result of the agent's
ignorance. On the other hand, this ignorance provides some conservatism,
reducing the noise due to early classifications if they are unreliable: in
practice we ignore the first two classifications made by  each volunteer, for
this reason. Whether or not an offline analysis out-performs an online one is a
matter for experiment.

@@

How fast does the \SW crowd classify subjects? Each software agent records the timestamp
of each classification its volunteer makes; by measuring the time lags between
successive classifications, we can make estimates of the crowd's classification
and contribution speed. We plot the former quantity for both \StageOne and
\StageTwo in \Fref{fig:results:speed}, normalizing the speed and time axes to
their respective totals. The fractional classification rates
in each stage of the survey fall off in approximately the same way, despite the
factor of nearly 20 difference in crowd size. \StageOne (consisting of
$\sim430,000$ subjects) was completed in around 5000 hours (by some 37,000
participants), with classification rates in the first few days averaging at
$\sim10^4$ per hour. This was stimulated by various forms of advertising (press
releases, and emails to registered Zooniverse users). The asymptotic
classification rate was  around ten times lower, at $\sim10^3$ per hour.
Interestingly, the average skill was found to be approximately constant over the
lifetime of \StageOne, leading to a contribution rate that tracks closely the
classification rate. (A 10\% increase in average skill was seen over the first
40 days, and a decrease of the same amount in the last 40 days -- not enough to
cause a significant difference between classification and contribution rate
behaviour, but perhaps reflecting a learning period at the start and then a
decrease
in participation later on.) \StageTwo (nearly 3400 subjects) was completed in around 600
hours by around 2000 participants, with classification rates starting at
$~\sim2000$ per hour and decaying to $\sim100$ per hour.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
\begin{center}
\caption{Total crowd and subject sample properties from the \cfhtls project.}
\label{tab:crowd:contributions}
\begin{tabular}{cccccccc}
  \hline
  \hline {Stage} & Subjects & Contribution                          & Agents & Skill      & Classifications          & Candidates & Information \\
                           & $\Ns$    & $\sum_k^{\Nv} \contribution_k$ (bits) & $\Nv$  & $\sum_k^{\Nv} \skill_k$ (bits) & $\sum_k^{\Nv} \thiseffort$ & $\Ncands$  & $\sum_j^{\Ns}\sum_k^{\Nv} \information_{j,k}$ (bits) \\
            \hline
                      1    & $427064$ & $1292016.3$ & $36982$ & $1471.9$ & $10802125$ & $3381$ & $91122.6$ \\
                      2    & $3679$   &   $21895.8$ &  $1964$ &  $102.4$ &   $224745$ &   $89$ &  $1640.4$ \\
  \hline \hline
\end{tabular}
\medskip\\
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $91-93\%$ and $<1\%$
(Stages I and II). This FPR translates to a purity of approximately 30\%.
We find an even
higher purity at \StageTwo (perhaps as a result of the least visible candidates already
being discarded): here, 100\% purity was reached at just under $90\%$
completeness. Because \SW is a supervised learning system, these numbers should
be taken to be upper limits to what we should expect for real lenses (which have
not been selected for a high visibility training set).
